Dependently Typed Multi-stage Calculus

### 2019-11-30 slides/main.tex

\begin{frame}[fragile]{What is Multi-stage Programming}
    A programming paradigm that enables
    \begin{itemize}
        \item Generation of code at run-time
        \item Evaluation of code at run-time
    \end{itemize}
    Applications
    \begin{itemize}
        \item Speeding up programs using run-time information \\
            (a.k.a. run-time specialization)
        \item DSLs
    \end{itemize}
    \note{
        So, what is multi-stage programming? It is a programming paradigm that
        enables generation of code at run-time and evaluation of code at
        run-time simultaneously. One of famous applications of multi-stage
        programming is speeding up programs using run-time information, which
        is also known as run-time specialization. A DSL is also a famous
        application such as a DSL for linear algebra or image processing.
    }
\end{frame}

@@

 # (run vadd3) [1;2] [4;5]
 [1;2] has type vec 2
 but was expected of type vec 3
 \end{Verbatim}
    \end{exampleblock}
    \note{
        Our approach to the problem is using dependent types. Dependent types
        are types dependent on values and can prevent illegal use of vadd3.
        Using dependent types, we can express the length of vectors in their
        types. For example, a vector [1;2] is as vec 2 and vadd3 is typed code
        of function from vec 3 type to vec 3. Now, if I apply wrong length
        vectors, [1;2] and [4;5] to vadd3, I get only type error not run-time
        error.
    }
\end{frame}

@@

\begin{frame}[fragile]{Kinding rules of \LMD}
    Kiding judgements also include stages.
    \begin{center}
        \( \G \vdash \tau :: K @ n \)
    \end{center}
    Because types may contains terms. (e.g. vec n)
    \begin{center}
        {\footnotesize
        \infrule[{\KApp}]{
            \G\V \sigma:: (\Pi x:\tau.K) @ n \andalso \G\V M:\tau @ n
        }{
            \G\V \sigma\ M::K[x\mapsto M] @ n
        } \\[2mm]
        \infrule[{\KAbs}]{
            \G\V \tau :: * @ n \andalso \G,x:\tau@ n \V \sigma::J@n
        }{
            \G\V(\Pi x:\tau.\sigma) :: (\Pi x:\tau.J) @ n
        }}
    \end{center}
    \note{
        Kiding judgements also include stages. Because types may contains terms
        such as n in vec n. It is natural that types dependent on a value live
        in the same stage of the value. So, application rule, K-App, says that
        if you want to apply term M to type sigma, they must live in the same
        stage n. K-Abs says that when sigma is kinded at stage n and variable x
        in the environment is on stage n also, you can abstract x using Pi.
    }
\end{frame}

>>>>

\begin{frame}[fragile]{What is Multi-stage Programming?}
    A programming paradigm that enables
    \begin{itemize}
        \item Generation of code at run-time
        \item Evaluation of code at run-time
    \end{itemize}
    Applications
    \begin{itemize}
        \item Speeding up programs using run-time information \\
            (a.k.a. run-time specialization)
        \item DSLs
    \end{itemize}
    \note{
        So, what is multi-stage programming? It is a programming paradigm that
        enables generation of code at run-time and evaluation of code at
        run-time simultaneously. One of famous applications of multi-stage
        programming is speeding up programs using run-time information, which
        is also known as run-time specialization. A DSL is also a famous
        application such as a DSL for linear algebra or image processing.
    }
\end{frame}

@@

 # (run vadd3) [1;2] [4;5]
 [1;2] has type vec 2
 but a value of type Vec 3 was expected
 \end{Verbatim}
    \end{exampleblock}
    \note{
        Our approach to the problem is using dependent types. Dependent types
        are types dependent on values and can prevent illegal use of vadd3.
        Using dependent types, we can express the length of vectors in their
        types. For example, a vector [1;2] is as vec 2 and vadd3 is typed code
        of function from vec 3 type to vec 3. Now, if I apply wrong length
        vectors, [1;2] and [4;5] to vadd3, I get only type error not run-time
        error.
    }
\end{frame}

@@

\begin{frame}[fragile]{Kinding rules of \LMD}
    Kiding judgements as well include stages.
    \begin{center}
        \( \G \vdash \tau :: K @ n \)
    \end{center}
    Because types may contains terms. (e.g. vec n)
    \begin{center}
        {\footnotesize
        \infrule[{\KApp}]{
            \G\V \sigma:: (\Pi x:\tau.K) @ n \andalso \G\V M:\tau @ n
        }{
            \G\V \sigma\ M::K[x\mapsto M] @ n
        } \\[2mm]
        \infrule[{\KAbs}]{
            \G\V \tau :: * @ n \andalso \G,x:\tau@ n \V \sigma::J@n
        }{
            \G\V(\Pi x:\tau.\sigma) :: (\Pi x:\tau.J) @ n
        }}
    \end{center}
    \note{
        Kiding judgements also include stages. Because types may contains terms
        such as n in vec n. It is natural that types dependent on a value live
        in the same stage of the value. So, application rule, K-App, says that
        if you want to apply term M to type sigma, they must live in the same
        stage n. K-Abs says that when sigma is kinded at stage n and variable x
        in the environment is on stage n also, you can abstract x using Pi.
    }
\end{frame}
